---
title: "2019 Pandalus Prey"
author: "Krista Kenyon"
date: "`r Sys.Date()`"
output:
  bookdown::word_document2:
    global_numbering: true
    fig_caption: true
  
---

<!---
Comments for the YAML header above (I can't place this comment above it)

If I want to customize settings of a component (e.g. output), place a ':' at the end.
  I couldn't realize why, say 'toc: true' kept producing errors despite being spaced
  appropriately. It turned out, I did not include the ':' after word_document2:

bookdown::word_document2 = produce a word document AND allow me to cross reference
  figures, tables, and headings.They are only hyperlinked in pdf and html outputs (not word)

NOT NEEDED WITH BOOKDOWN -> number_sections: true = automatically numbering the sections.

global_numbering: true -> word and html outputs label Figures as Fig 1, 2, etc and sections as 1.0, 1.1 etc. This automatically happens with pdf outputs but not the other two.

** At end add in -> 'toc: true'. This adds a  table of contents at the beginning of the document

fig_caption: true = render figures with captions

One user also used 'bibliography' (main function like 'output') to insert which 
  files he had his citations stored. He included 2 files and could reference them.
  (didn't show how). https://www.youtube.com/watch?v=YlcK9Ef31KE

?rmarkdown::word_document= = shows me all the formatting options 

***** I would LOVE to learn how to direct where the output documents are saved.
  I know there are ways.... I just can't find them right now
  
***** For formating pdf text size, fonts, line spacing or insert line numbers etc look at https://www.youtube.com/watch?v=kC6JFcM5e9M
--->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 

# Below are the packages needed for this markdown document
       # The 'include=FALSE' parameter means this r chunk won't be shown in the report

# library(ggplot2) 
# library(tidyverse)


### Notes on cross-referencing figures and tables ###

# In the YAML Header, the 'bookdown::word_document2' parameter gives the ability to cross-reference
# I can reference in text using Figure \@ref(fig:label). It hyperlinks it in pdf but not word. It also gives different Figure number in pdf and word. So there are still things to figure out.
  # to cross reference the figure needs a unique code-chunk name, and have a figure caption

# Code chunk names should be one word or separated with '-', not with spaces, '_', or '.'
  #{r code-chunk}
# I can create figure captions by including 'fig.cap=' in the {r code-chunk, fig.cap=caption text}.

  # If text \@ref(fig:label) is included in notes in the R markdown portion, there will be an error that occurs when Knitting. Therefore included here.
```

<!--
This will form the template for the annual reports presenting the raw fish stomach content species data.
-->

<!--
General formatting notes unique to Rmarkdown.

- If you want text to be on separate lines, you NEED to include 2 spaces at the end of the proceeding line.

- There needs to be a line in between text and a list in order for the formatting to create the list. This does create a space between them in the output as well. I have not looked up if there is a way to adjust the output spacing.
-->

**The following are the pre-amble pages that will need to be included at some point.**

- Document Title Page  <!--- look here https://www.youtube.com/watch?v=kC6JFcM5e9M --->
- Canadian Data Report of Fisheries and Aquatic Sciences (smaller heading)  
- Report Title Page  
- Page of Ownership and How to Cite Report  
- Table of Contents  
- List of Tables  
- List of Figures   

One way to create the Table of Contents is the 'settings function' to the right of 'Knit' -> Output Options -> check Include TOC. This will need to be formatted though to fit the style of the DFO documents.  

<!--- \newpage creates a page break --->
\newpage


# Abstract

# Introduction

This section will include Figure 1
<!--
Figure 1: jpeg of different fishing zones
-->


# Methods

This section will include Figure 2 
<!--
Figure 2: jpeg of the different stomach sampling locations. Looks like made in ArcGIS
-->

This section will include Table 1
<!--
Table 1: # of stomach sampling locations in each zone per depth stratum.

Data that I will need include sample location depths in each stratum.
-->


## Collection of Stomach Samples

## Stomach Content Analysis

## Data Analysis

# Results

```{r import_data, include=FALSE}

### THIS IS ASSUMING THAT DATASET IS ALREADY SUBSETTED FOR YEAR, PREDATOR SPECIES, AND SURVEY AREA ###

### import data
stomach <- read.csv('data/raw/2013_test_stomach_data.csv') 
  # assumes header=TRUE and knows it's separated by commas


### Adding count of stomachs sampled per fish. Some figures will need this data

stomach$count.stomach <- 1 
  # add new column 'count.stomach' to give a numerical value for the number of stomachs sampled from each fish. Will be necessary when creating plots.


### Adding a column with predatory 'category' names (i.e. skate, redfish, Greenland halibut, Atlantic cod) based on species code (column pred.sp)

stomach$pred.name = stomach$pred.sp  
  # creates new column that is duplicate of pred.sp, and renames to pred. name
library(dplyr)

stomach <- stomach %>% 
  mutate(pred.name=recode(pred.name, "438"="Atlantic cod", "892"="Greenland halibut", "793" ="redfish","794"="redfish", "90"="skate"))  
  # replaces species codes with predatory categories. 
  # currently the above code only includes the species codes for 2013

### Possible species codes for predator species categories
  ## Greenland halibut = 892
  ## Atlantic cod = 438
  ## Redfish = 792, 793, 794, 997, 998
  ## Skates = 80, 88, 89, 90, 91, 92, 94, 95, 96, 97
  ## See 'Predator Codes and Names.txt' file in 'Scripts' folder for list of species names attached to each code


### Checking our modifications to the dataframe

which(is.na(stomach))
  # checking if there are blank values or NA's in the dataframe

str(stomach) 
  # checking new dataframe structure to ensure all updates are good

### str(stomach) return 
# Show in New Window
# 'data.frame':	2300 obs. of  18 variables:
#  $ content.key  : chr  "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" ...
#  $ prey.sp      : int  9998 9998 9998 9998 4770 4770 4770 4770 9998 9998 ...
#  $ weight       : num  0 0 0 0 0.0015 0.0015 0.0015 0.0015 0 0 ...
#  $ count        : int  0 0 0 0 1 1 1 1 0 0 ...
#  $ pred.sp      : int  892 892 892 892 892 892 892 892 892 892 ...
#  $ length       : int  40 40 40 40 45 45 45 45 52 52 ...
#  $ year         : int  2013 2013 2013 2013 2013 2013 2013 2013 2013 2013 ...
#  $ start.depth  : int  823 823 823 823 823 823 823 823 823 823 ...
#  $ end.depth    : int  829 829 829 829 829 829 829 829 829 829 ...
#  $ lat.deg      : int  64 64 64 64 64 64 64 64 64 64 ...
#  $ lat.min      : num  27.9 27.9 27.9 27.9 27.9 ...
#  $ long.deg     : int  57 57 57 57 57 57 57 57 57 57 ...
#  $ long.min     : num  59.8 59.8 59.8 59.8 59.8 ...
#  $ sa.code      : chr  "2G" "RISA" "SFA2EX" "SFA3" ...
#  $ sort.key     : int  10 7 6 8 10 7 6 8 10 7 ...
#  $ trawl.id     : int  72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 ...
#  $ count.stomach: num  1 1 1 1 1 1 1 1 1 1 ...
#  $ pred.name    : chr  "Greenland halibut" "Greenland halibut" "Greenland halibut" "Greenland halibut" 
```


## Predator Distribution by Assessment Area, Length, and Depth

This section will include Figure 3. 

It will contain cool text that will reference Figure \@ref(fig:pie-chart)
<!--
Figure 3: pie chart showing proportion of fish species caught

Data I will need is simply # of fish caught per species
-->

```{r pie-chart, echo=FALSE, fig.cap="*Proportion of Atlantic Cod (*Gadus morhua*), Greenland Halibut (*Reinhardtius hippoglossides*), redfish (*Sebastes sp.*), and skate (*Rajidae*) stomachs analyzed, calculated as a percent total.*"}


### Creating new dataframe detailing total sum of stomachs sampled per predatory category

predsp.total <- aggregate(stomach$count.stomach, list(stomach$pred.name), sum)
  ## aggregate sum of number of predator stomachs sampled

# predsp.total  #    !!!! Make it a 'comment' before knitting !!!! otherwise table enters output
  # checking dataframe. 

### Renaming the columns in the new dataframe
names(predsp.total)[names(predsp.total)=="Group.1"] <- "pred.name"
names(predsp.total)[names(predsp.total)=="x"] <- "sum.count.stomach"
# names(predsp.total)  #  !!!! Make it a 'comment' before knitting !!!! 

## I tried the below code to rename with 'dyplr' package but it wasn't working
# predsp.total %>% 
#   rename(
#     pred.name = Group.1, 
#     sum.count.stomach = x
#     )
# str(predsp.total)

#str(predsp.total)    #  !!!! Make it a 'comment' before knitting !!!!
  ## saving dataframe structure
# 'data.frame':	4 obs. of  2 variables:
#  $ pred.name        : chr  "Atlantic cod" "Greenland halibut" "redfish" "skate"
#  $ sum.count.stomach: num  20 1340 900 40


### Creating the Pie Chart

pie(predsp.total$sum.count.stomach, labels=predsp.total$pred.name)


### Formatting the Pie Chart ######################## STILL NEED TO DO ###################3

# This will be completed once I have the 2019 data.


```


This section will include Figure 4 
<!--
Figure 4: bar chart of fish caught per region.

Data I will need is # fish caught per species per region
-->

This section will include Figure 5  

<!--
Figure 5: 4-paneled bar chart of number of stomach's sampled from different fish caught at each depth stratum. Each panel shows different fish species. 

Data I will need is # fish caught per depth stratum per species
-->

## Proportion of Full vs. Empty Stomachs

This section will include Table 2
<!--
Table 2: Table showing total fish with full and empty stomachs for each fish class size for each fish species

Data I will need includes length (cm) of caught fish, if stomach was full or empty, for each fish species.
-->

## Composition of Prey Items

### Composition of Prey Items by %W

### Composition of Prey Items by %N

This section will include Table 3  

<!--
Table 3: Table showing percent weight and percent number of each prey species taxa within the Atlantic cod stomachs. 

Data I will need includes weight and number of prey species within Atlantic cod stomachs
-->

This section will include Table 4  

<!--
Table 4: Table showing percent weight and percent number of each prey species taxa within the Greenland Halibut stomachs. 

Data I will need includes weight and number of prey species within Atlantic cod stomachs
-->

This section will include Table 5

<!--
Table 5: Table showing percent weight and percent number of each prey species taxa within the redfish stomachs. 

Data I will need includes weight and number of prey species within redfish stomachs
-->

This section will include Table 6  

<!--
Table 6: Table showing percent weight and percent number of each prey species taxa within the skates stomachs. 

Data I will need includes weight and number of prey species within skates stomachs
-->

## Pandalus as Prey

### Relative Proportion of Pandalid Prey by Predatory Species

This section will include Figure 6

<!--
Figure 6: 2-paneled bar chart of percent weight and percent number of different shrimp species within the stomachs of the four fish species.  

Data I will need is weight and number of shrimp in each target fish species
-->

### Relative Proportion of Pandalid Prey by Survey Area and Depth

This section will include Figure 7

<!--
Figure 7: 2-paneled bar chart of percent weight and percent number of pandalus shrimp species (combined) within all stomachs by the different depth stratum by fishing area  

Data I will need is weight and number of pandalus shrimp, depth stratum fish was caught, and fishing region that fish were caught
-->

### Relative Proportion of Pandalid Prey by Size Class for Atlantic Cod (*Gadus morhua*) 

This section will include Figure 8

<!--
Figure 8: 2-paneled bar chart of percent weight and percent number different shrimp species within Atlantic cod 

Data I will need is weight and number of shrimp species in Atlantic cod stomachs. 
-->

### Relative Proportion of Pandalid Prey by Size Class for Greenland Halibut (*Reinhardtius hippoglossides*)

This section will include Figure 9

<!--
Figure 9: 2-paneled bar chart of percent weight and percent number different shrimp species within Greenland halibut 

Data I will need is weight and number of shrimp species in Greenland halibut stomachs. 
-->

### Relative Proportion of Pandalid Prey by Size Class for Redfishes (*Sebastes sp.*)

### Relative Proportion of Pandalid Prey by Size Class for Skates (*Rajidae*)

This section will include Figure 10

<!--
Figure 10: 2-paneled bar chart of percent weight and percent number different shrimp species within skate

Data I will need is weight and number of shrimp species in skate stomachs. 
-->

# Acknowledgements

# References
