---
title: "2019 Pandalus Prey"
author: "Krista Kenyon"
date: "`r Sys.Date()`"
output:
  bookdown::word_document2:
    global_numbering: true
    fig_caption: true
  bookdown::pdf_document2:
bibliography: references.bib
csl: the-journal-of-wildlife-management.csl
  
---

<!---
Comments for the YAML header above (I can't place this comment above it)

If I want to customize settings of a component (e.g. output), place a ':' at the end.
  I couldn't realize why, say 'toc: true' kept producing errors despite being spaced
  appropriately. It turned out, I did not include the ':' after word_document2:

bookdown::word_document2 = produce a word document AND allow me to cross reference
  figures, tables, and headings.They are only hyperlinked in pdf and html outputs (not word)

NOT NEEDED WITH BOOKDOWN -> number_sections: true = automatically numbering the sections.

global_numbering: true -> word and html outputs label Figures as Fig 1, 2, etc and sections as 1.0, 1.1 etc. This automatically happens with pdf outputs but not the other two.

** At end add in -> 'toc: true'. This adds a  table of contents at the beginning of the document

fig_caption: true = render figures with captions

One user also used 'bibliography' (main function like 'output') to insert which 
  files he had his citations stored. He included 2 files and could reference them.
  (didn't show how). https://www.youtube.com/watch?v=YlcK9Ef31KE

?rmarkdown::word_document= = shows me all the formatting options 

***** I would LOVE to learn how to direct where the output documents are saved.
  I know there are ways.... I just can't find them right now
  
***** For formating pdf text size, fonts, line spacing or insert line numbers etc look at https://www.youtube.com/watch?v=kC6JFcM5e9M
--->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 

# This code chunk sets the parameters for the document code chunks

# The 'include=FALSE' parameter means this particular r chunk won't be shown in the report


######################### Required Packages ####################################


library(ggplot2)
library(tidyverse)
library(dplyr)
library(ggrepel)           # Fig 3: ensures labels don't overlap
library(reshape2)          # Fig 4
library(viridis)                 # Fig 5 - color blind friendly palette
library(gt)                      # Table formatting
library(gtExtras)                # Additional table formatting options
library(flextable)        # Table 1 options
library(kableExtra)       # Table 1 extra Kable options that may only work in HTML
# library(officer)          # helps flextable formatting transfer to Word doc


######## Setting up the Theme

# At the 5 Minute mark it goes into establishing a universal theme for all charts https://www.youtube.com/watch?v=4KMrOZ_BbKE
    # not sure it would work here because it has charts and tables??

# examples from video is
# theme_set(theme_bw() + theme(title = element_text(colour = "steelblue", face = "bold")))


################# Notes on cross-referencing figures and tables ################


# In the YAML Header, the 'bookdown::word_document2' parameter gives the ability to cross-reference

# I can reference in text using Figure \@ref(fig:label). It hyperlinks it in pdf but not word. It also gives different Figure number in pdf and word. So there are still things to figure out.
  # to cross reference the figure needs a unique code-chunk name, and have a figure caption

# Code chunk names should be one word or separated with '-', not with spaces, '_', or '.'
  #{r code-chunk}

# I can create figure captions by including 'fig.cap=' in the {r code-chunk, fig.cap=caption text}.

  # If text \@ref(fig:label) is included in notes in the R markdown portion, there will be an error that occurs when Knitting. Therefore included here.
```

<!--
This will form the template for the annual reports presenting the raw fish stomach content species data.
-->

<!--
General formatting notes unique to Rmarkdown.

- If you want text to be on separate lines, you NEED to include 2 spaces at the end of the proceeding line.

- There needs to be a line in between text and a list in order for the formatting to create the list. This does create a space between them in the output as well. I have not looked up if there is a way to adjust the output spacing.
-->

**The following are the pre-amble pages that will need to be included at some point.**

- Document Title Page  <!--- look here https://www.youtube.com/watch?v=kC6JFcM5e9M --->
- Canadian Data Report of Fisheries and Aquatic Sciences (smaller heading)  
- Report Title Page  
- Page of Ownership and How to Cite Report  
- Table of Contents  
- List of Tables  
- List of Figures   

One way to create the Table of Contents is the 'settings function' to the right of 'Knit' -> Output Options -> check Include TOC. This will need to be formatted though to fit the style of the DFO documents.  

<!--- \newpage creates a page break --->
\newpage


# Abstract



```{r import_data, include=FALSE}

# This code chunk:
    # imports the data
    # manipulates the data it to create the base prey dataframe
    # manipulates the data it to create the base predator dataframe
    # checks & saves the dataframe structures

# The data imported must be structured consistently in order to utilize this coding.

# All code to check dataframe structure are commented out. If not they will produce outputs into the report.


############################ Key Data Structure Assumption #####################

# It is assumed that the imported data will initially have a structure consistent to the example saved immediately below. Particularly the variable names and variable data type (e.g. chr, int, num)



### Example initial imported dataset structure 

# 'data.frame':	2300 obs. of  16 variables:
#  $ content.key: chr  "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" ...
#  $ prey.sp    : int  9998 9998 9998 9998 4770 4770 4770 4770 9998 9998 ...
#  $ weight     : num  0 0 0 0 0.0015 0.0015 0.0015 0.0015 0 0 ...
#  $ count      : int  0 0 0 0 1 1 1 1 0 0 ...
#  $ pred.sp    : int  892 892 892 892 892 892 892 892 892 892 ...
#  $ length     : int  40 40 40 40 45 45 45 45 52 52 ...
#  $ year       : int  2013 2013 2013 2013 2013 2013 2013 2013 2013 2013 ...
#  $ start.depth: int  823 823 823 823 823 823 823 823 823 823 ...
#  $ end.depth  : int  829 829 829 829 829 829 829 829 829 829 ...
#  $ lat.deg    : int  64 64 64 64 64 64 64 64 64 64 ...
#  $ lat.min    : num  27.9 27.9 27.9 27.9 27.9 ...
#  $ long.deg   : int  57 57 57 57 57 57 57 57 57 57 ...
#  $ long.min   : num  59.8 59.8 59.8 59.8 59.8 ...
#  $ sa.code    : chr  "2G" "RISA" "SFA2EX" "SFA3" ...
#  $ sort.key   : int  10 7 6 8 10 7 6 8 10 7 ...
#  $ trawl.id   : int  72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 ...


####################### Importing Your Data ####################################


### import data

prey <- read.csv('data/raw/2019_StomachContent_test.csv') 
    # assumes header=TRUE and knows it's separated by commas

str(prey)
    # check initial data structure. Does it match example above?


########################## Creating Base Prey Dataframe ########################


# This code chunk:
    # manipulates the data it to create the base prey dataframe
    # checks & saves the dataframe structure

# 'prey' dataframe: each row is a new prey item per sampled stomachs

# Dataframe transformations:
    # add column with shrimp fishing regions (EAZ, WAZ, SFA4)
    # add column with predator 'category' names (Atlantic code, Greenland halibut, Redfish, Skate)
    # add column with depth range category (100-200 m, 200-300 m, 300-400 m, 400-500 m, 500-750 m)
    # add predator length range category (binned into 5 cm length groups)

# Finish with checking and storing dataframe structure



### Adding 'region' category (WAZ, EAZ, SFA4) 

prey$region = prey$Study.Area  
    # creates new column that is duplicate of sa.code, and renames to pred. name

prey <- prey %>% 
  mutate(region=recode(region, "RISA"="EAZ", "SFA2EX"="EAZ", "SFA3" = "WAZ",
                       "2G" = "SFA4"))
    # renaming values in the prey$region column



### Adding predator category

prey$pred.name = prey$OS_ID  
    # creates new column that is duplicate of pred.sp, and renames to pred. name

prey <- prey %>%
  mutate(pred.name=recode(pred.name, "892" = "Greenland halibut",
                          "438" = "Atlantic cod",
                          "792" = "Redfish", "793" = "Redfish", "794" = "Redfish",
                          "997" = "Redfish", "998" = "Redfish",
                          "80" = "Skate", "88" = "Skate", "89" = "Skate",
                          "90" = "Skate", "91" = "Skate", "92" = "Skate",
                          "94" = "Skate", "95" = "Skate", "96" = "Skate",
                          "97" = "Skate"))
    # replaces species codes with predatory categories in new column. 
    # See 'Predator Codes and Names.txt' file in 'Scripts' folder for list of species names attached to each code



### Adding depth range category

# The Access database does have a depth-range column BUT it has with a variety of overlapping ranges.
# Below code creates desired depth ranges based on start-end depths
# This ensures code is repeatable even if new depth ranges are added into the database

prey <- prey %>% 
  mutate(depth = case_when(Start.Depth >= 100 & End.Depth >= 100 & 
                             Start.Depth <= 200 & End.Depth <= 200 ~ "100-200 m",
                           
                           Start.Depth >= 200 & End.Depth >= 200 & 
                             Start.Depth <= 300 & End.Depth <= 300 ~ "200-300 m",
                           
                           Start.Depth >= 300 & End.Depth >= 300 & 
                             Start.Depth <= 400 & End.Depth <= 400 ~ "300-400 m",
                           
                           Start.Depth >= 400 & End.Depth >= 400 & 
                             Start.Depth <= 500 & End.Depth <= 500 ~ "400-500 m",
                           
                           Start.Depth >= 500 & End.Depth >= 500 & 
                             Start.Depth <= 750 & End.Depth <= 750 ~ "500-750 m"))
    # yes multiple categories both use the same 'start/end' points.
    # the combination of requirements within each depth category means there isn't actually overlap



### Adding predator length range category

# Binning the data
prey <- prey %>% 
  mutate(length.range = cut(Length, breaks=seq(from = 0, to = 5.99, by = 5.99))) %>% 
    # sets first bin as < 6
  
  mutate(length.range = cut(Length, breaks=seq(from = 5.99, to = max(Length)+5, by = 5)))   
    # sets remaining bins as # >=6 and <11 (or range 6-10). Etc
    # max(Length) ends the sequence at the length of the longest fish sampled
    # longest fish ends up being excluded with this
    # max(Length)+5 ensures the longest fish will be included


# Reformatting the binned outputs 
prey <- prey %>%
  mutate(length.range = str_replace(length.range, ',', '-')) %>%        
    # replaces , with -
  
  mutate(length.range = gsub("[(]|[]]", '', length.range,)) %>%        
    # removes ( or ]. Equivalent code wasn't working with str_replace
  
  mutate(length.range = str_replace(length.range, "5.99", "6"))  %>%    
    # replaces 5.99 with 6
  
  mutate(length.range = str_replace(length.range, "1$", "0")) %>%       
    # if the last digit is 1, it is replaced with a 0
  
  mutate(length.range = str_replace(length.range, "6$", "5"))           
    # if the last digit is 6, it is replaced with a 5



### Any missing values?

colSums(is.na(prey)) 
    # check if there are any NAs remaining (shows # NA's per column)



### Saving the prey dataframe structure

str(prey) 

# 'data.frame':	1236 obs. of  23 variables:
 
# $ ContentKey        : chr  "62114003_892_3_1_-14_1733" "62114003_892_3_1_-15_1732" "62114003_892_3_1_-16_1728"...
# $ Prey_OS_ID        : int  6967 9998 6967 6967 6967 6967 8020 8020 4950 8530 ...
# $ PreyWt            : num  0.3 NA 0.286 0.091 0.079 ...
# $ Prey_Count        : int  NA NA NA NA NA NA NA NA NA NA ...
# $ OS_ID             : int  892 892 892 892 892 892 892 892 892 892 ...
# $ Length            : num  24.5 20 20 20.5 17 14 37.5 27 27 33 ...
# $ Trawl_ID          : int  62114003 62114003 62114003 62114003 62114003 62114003 62114003 62114003 62114003 ...
# $ Year              : int  2019 2019 2019 2019 2019 2019 2019 2019 2019 2019 ...
# $ Start.Depth       : int  269 269 269 269 269 269 269 269 269 269 ...
# $ End.Depth         : int  271 271 271 271 271 271 271 271 271 271 ...
# $ Depth.Range       : chr  "201-300" "201-300" "201-300" "201-300" ...
# $ Study.Area        : chr  "2G" "2G" "2G" "2G" ...
# $ Scientific.Name   : chr  "REINHARDTIUS HIPPOGLOSSOIDES" "REINHARDTIUS HIPPOGLOSSOIDES" ...
# $ CommonName        : chr  "GREENLAND HALIBUT, TURBOT" "GREENLAND HALIBUT, TURBOT" ...
# $ Start.Lat.Degree  : int  57 57 57 57 57 57 57 57 57 57 ...
# $ Start.Lat.Minutes : num  45.1 45.1 45.1 45.1 45.1 ...
# $ Start.Long.Degree : int  60 60 60 60 60 60 60 60 60 60 ...
# $ Start.Long.Minutes: num  24.2 24.2 24.2 24.2 24.2 ...
# $ FishKey           : chr  "62114003_892_3_1_-14" "62114003_892_3_1_-15" "62114003_892_3_1_-16" ...
# $ region            : chr  "SFA4" "SFA4" "SFA4" "SFA4" ...
# $ pred.name         : chr  "Greenland halibut" "Greenland halibut" "Greenland halibut" "Greenland halibut" ...
# $ depth             : chr  "200-300 m" "200-300 m" "200-300 m" "200-300 m" ...
# $ length.range      : chr  "21-25" "16-20" "16-20" "16-20" ...


###################### Creating Base Predator Dataframe ########################


# This code chunk:
    # manipulates the data it to create the base predator dataframe
    # checks & saves the dataframe structure

# 'pred' dataframe: each row is a new fish that's stomach was sampled 

# Dataframe transformations:
    # check number of fish sampled
    # duplicate prey dataframe
    # remove duplicate fish-id values (from multiple prey found within stomachs)
    # confirm number of sampled fish retained is correct
    # add column for number of stomachs sampled

# Finish with checking and storing dataframe structure



### checking how many unique fish were sampled

length(unique(prey$FishKey))
    # 2019 data: 644 observations



### duplicating base prey dataframe

pred <- prey



### Retain one row per sampled fish

# removing duplicated values for FishKey so that there is one entry per predator
pred <- pred[!duplicated(pred$FishKey),]
    # removes duplicates in FishKey

# CHECK NUMBER VARIABLES! Does it match # unique fish sampled above

 str(pred) 
    # for 2019 data: 644 observations



### Adding category for the number of stomachs sampled per fish

pred$sampled.stomach <- 1 



### Saving Structure of Base Dataframe

# str(pred)

# 'data.frame':	644 obs. of  22 variables:

# $ ContentKey        : chr  "62114003_892_3_1_-14_1733" "62114003_892_3_1_-15_1732" ...
# $ Prey_OS_ID        : int  6967 9998 6967 6967 6967 6967 8020 8020 8530 8020 ...
# $ PreyWt            : num  0.3 NA 0.286 0.091 0.079 ...
# $ Prey_Count        : int  NA NA NA NA NA NA NA NA NA NA ...
# $ OS_ID             : int  892 892 892 892 892 892 892 892 892 892 ...
# $ Length            : num  24.5 20 20 20.5 17 14 37.5 27 33 33.5 ...
# $ Trawl_ID          : int  62114003 62114003 62114003 62114003 62114003 62114003 62114003 62114003 62114003 ...
# $ Year              : int  2019 2019 2019 2019 2019 2019 2019 2019 2019 2019 ...
# $ Start.Depth       : int  269 269 269 269 269 269 269 269 269 269 ...
# $ End.Depth         : int  271 271 271 271 271 271 271 271 271 271 ...
# $ Depth.Range       : chr  "201-300" "201-300" "201-300" "201-300" ...
# $ Study.Area        : chr  "2G" "2G" "2G" "2G" ...
# $ Scientific.Name   : chr  "REINHARDTIUS HIPPOGLOSSOIDES" "REINHARDTIUS HIPPOGLOSSOIDES" ...
# $ CommonName        : chr  "GREENLAND HALIBUT, TURBOT" "GREENLAND HALIBUT, TURBOT" ...
# $ Start.Lat.Degree  : int  57 57 57 57 57 57 57 57 57 57 ...
# $ Start.Lat.Minutes : num  45.1 45.1 45.1 45.1 45.1 ...
# $ Start.Long.Degree : int  60 60 60 60 60 60 60 60 60 60 ...
# $ Start.Long.Minutes: num  24.2 24.2 24.2 24.2 24.2 ...
# $ FishKey           : chr  "62114003_892_3_1_-14" "62114003_892_3_1_-15" "62114003_892_3_1_-16" ...
# $ region            : chr  "SFA4" "SFA4" "SFA4" "SFA4" ...
# $ pred.name         : chr  "Greenland halibut" "Greenland halibut" "Greenland halibut" "Greenland halibut" ...
# $ sampled.stomach      : num  1 1 1 1 1 1 1 1 1 1 ...

```

# Introduction

Testing out references [@polaczek2023]

This section will include Figure 1
<!--
Figure 1: jpeg of different fishing zones
-->


# Methods

This section will include Figure 2 
<!--
Figure 2: jpeg of the different stomach sampling locations. Looks like made in ArcGIS
-->

This section will include Table 1.
<!--
Table 1: # of stomach sampling locations in each zone per depth stratum.

Data that I will need include sample location depths in each stratum.
-->


Piping external code into Markdown via 'file =' for Table\@ref(tab:sample-locations)

```{r sample-locations, file = 'scripts/Script_Table01_Sample-Locations.R', echo=FALSE}

# before knitting ensure to comment out in source code:
    # loading libraries
    # importing prey base dataframes (below loading libraries)
    # importing table dataframe (right before creating the table itself)
    # code that produces console outputs except for the final table
        # e.g. str()

# before knitting update the year on any source code outputs:
    # there are 2 'write.csv' codes

```





## Collection of Stomach Samples

## Stomach Content Analysis

## Data Analysis

# Results


## Predator Distribution by Assessment Area, Length, and Depth

This section will include Figure 3. 

It will contain cool text that will reference Figure \@ref(fig:pie-chart)
<!--
Figure 3: pie chart showing proportion of fish species caught

Data I will need is simply # of fish caught per species
-->

```{r pie-chart, echo=FALSE, fig.cap="*Proportion of Atlantic Cod (*Gadus morhua*), Greenland Halibut (*Reinhardtius hippoglossides*), redfish (*Sebastes sp.*), and skate (*Rajidae*) stomachs analyzed, calculated as a percent total.*"}

# The purpose of this code chunk is to create Figure 3

# Figure Caption is currently from Polaczek 2023

# The sections of this code chunk are:
    # Creating a new dataframe for Figure 3
    # Creating and Formatting Figure 3 pie chart
    # Recording Additional Geom_Label arguments that we may want to include in further edits 

# Tutorial and experimental codes for Fig 3 are saved in (scripts/draft code/Draft_Script_Fig03_PieChart.R).

# Those notes include other pie chart packages, functions, and arguments that I learned but did not end up using

# Additional notes on geom_label and geom_label_repel are saved at the bottom of (Script_Figure3_PieChart.R)


#################### Creating new dataframe for Figure #########################


# This dataframe will detail the total sum of stomachs sampled per predator category

# All code to check dataframe structure are commented out. If not they will produce outputs into the report.

# library(ggrepel) 
    # included in code chunk 'r setup'



### Create figure dataframe

pred.total <- aggregate(pred$sampled.stomach, list(pred$pred.name), sum)

# pred.total 



### Renaming the columns in the new dataframe

names(pred.total)[names(pred.total) == "Group.1"] <- "pred.name"

names(pred.total)[names(pred.total) == "x"] <- "stomach.total"

# names(pred.total)



### Creating figure labels

pred.total$label <- paste(pred.total$stomach.total, sep = "
", 
                          paste0( "(", round(100 * pred.total$stomach.total / sum(pred.total$stomach.total), 0 ), "%)" ))
    # Label structure will be 'total stomach sampled (% of stomachs)'
    # DO NOT ADJUST THE SPACING AFTER [sep = "..."]. This splits the values into 2 lines. Adding spaces will uncenter second line. 
    # paste() puts spaces between labels
    # paste0() does not have spaces between labels. Which means there is no space between # and % (e.g. 40% vs 40 %)



### Checking Final Database Structure

# str(pred.total)


### Saving Final Database Structure 

# 'data.frame':	4 obs. of  3 variables:
#  $ pred.name    : chr  "Atlantic cod" "Greenland halibut" "Redfish" "Skate"
#  $ stomach.total: num  20 1340 900 40
#  $ label        : chr  "20\n(1%)" "1340\n(58%)" "900\n(39%)" "40\n(2%)"


################### Building and Formatting the Pie Chart ######################


### Creating the pie chart

ggplot(pred.total, aes(x = "", y = stomach.total, fill = pred.name)) +
  
  coord_polar(theta = "y") + 
        # basing the circular coordinates on the y value (stomach.total)
  
  geom_col(color = "white", linewidth = 0.5) + 
       # creates white line separations between pieces
  
  theme_void() + 
        # removes the default theme (grey background & external grid lines)
  
  geom_label(aes(x = 1.65, label = label),  
        # x=# moves labels along 'x-axis'. Goal here is outside the pie chart.
        # label = pred.total$label created above. 
        # with 'aes()' we don't need 'pred.total$..'. Knows we are in 
    size = 2.5, 
        # adjusts size of label text
    position = position_stack(vjus = 0.5), 
        # defines 'y-axis' position of label. 'vjus=0.5' is centered.
    show.legend = FALSE) +
        # Removes geom_label default of adding an 'a' to legend colors
  
  theme(
    legend.title = element_blank(), 
        # removes legend title
    legend.key.size = unit(3, "mm"), 
        # adjusts width of legend color symbols
    legend.key.height = unit(3, "mm"), 
        # adjusts height of legend color symbols
    legend.key.spacing.y = unit(1, "mm"),
        # defines space between legend items
  )

```


This section will include Figure 4 

It will contain cool text that will reference Figure \@ref(fig:samples-zone)
<!--
Figure 4: bar chart of fish caught per region.

Data I will need is # fish caught per species per region
-->

```{r samples-zone, echo=FALSE, fig.cap="*Number of Atlantic Cod (*Gadus morhua*), Greenland Halibut (*Reinhardtius hippoglossides*), redfish (*Sebastes sp.*), and skate (*Rajidae*) stomachs analyzed from the Eastern Assessment Zone (EAZ), the Western Assessment Zone (WAZ), and Shrimp Fishing Area 4 (SFA 4)*"}

# The purpose of this code chunk is to create Figure 4

# Figure Caption is currently from Polaczek 2023

# The sections of this code chunk are:
    # Creating a new dataframe for Figure 4
    # Creating and Formatting Figure 4 bar chart

# If desired, additional notes and links to tutorial youtubes/webpages I used to create Figure 4 are recorded in (scripts/draft code/Draft_Script_Fig4_Bar-sample-region.R)

# Those notes include sites and tutorials that assisted me with:
    # Creating figure dataset
    # GGPLOT2 Code for Grouped Pie Charts
    # additional ggplot2 arguments not used in this chart



################### Creating New Dataframe for Fig 4 ###########################

# This dataframe will detail the total sum of stomachs sampled per predator category in each region

# All code to check dataframe structure are commented out. If not they will produce outputs into the report.

# New Dataframe structure will need:
    # Two categorical variables:
        # x-axis: regions
        # fill: predator category
    # One numerical variable: 
        # y-axis: total samples per predator species per region

# library(reshape2) 
    # included in code chunk 'r setup'



### create Fig 4 dataframe: sum stomach samples per species, per region

wide.data <- dcast(pred, region~pred.name, value.var="sampled.stomach", sum)
    # creates a dataframe in a 'wide format' and we will want a 'long format' dataframe
    # Tutorial 1 in Fig 4 draft code doc

# wide.data



### converting 'wide format' to 'long format'

region.total <- melt(data = wide.data,                       
      # melt() converts from wide (sp. spread across rows) to long format (sp. compiled into one row)
                     id.vars = "region",                     
      # id.vars : vector of variables I want to stack
                     measured.vars = c("Atlantic cod",       
                                       "Greenland halibut", 
                                       "Redfish", "Skate"),
      # measured.vars : identifies the columns that have the data I want to stack
                     variable.name = "pred.name",            
      # variable.name : what I want to call the stacked column
                     value.name = "stomach.sum")             
      # value.name : name of the column of my stacked measurements in my output data frame
    
      # Tutorial 2 in Fig 4 draft code doc melt()
    

# region.total



### Checking Final Database Structure

# str(region.total)



### Saving Final Database Structure 

# 'data.frame':	12 obs. of  3 variables:
# $ region     : chr  "EAZ" "SFA4" "WAZ" "EAZ" ...
# $ pred.name  : Factor w/ 4 levels "Atlantic cod",..: 1 1 1 2 2 2 3 3 3 4 ...
# $ stomach.sum: num  2 1 0 198 92 43 34 28 3 120 ...


############### Building and Formatting the Grouped Bar Chart ##################


### Grouped Bar Chart

ggplot(data = region.total,
       aes(fill=pred.name, y=stomach.sum, x=region)) + 
  
  geom_bar(position = position_dodge(),                          
         # makes the bars beside each other while being compatible with additional arguments
           width = 0.75,                                         
         # defines width of bars
           stat="identity",
           color = "black") +                                    
         # makes black boarders around bars
  
  geom_text(aes(label = stomach.sum),                            
         # label bar charts based on y value
            vjust = -0.5,                                        
         # puts labels above bars
            size = 3,                                            
         # changes size of labels
            position = position_dodge(width = 0.75)) +
         # matches the labels with our grouped bars (match widths in geom_text and geom_bar)
  
  scale_fill_manual(values = c("Atlantic cod" = "#FCFED4",       
                               "Greenland halibut" = "#CCEDB1", 
                               "Redfish" = "#41B7C4", 
                               "Skate" ="#FF9999" )) +
        # above assigns colour to particular predator categories
  
  theme_minimal() +                                              
        # minimal theme removes right/left boarders
        # minimal theme places x-axis directly below bars
        # minimal theme removes gray background color
  
  theme(legend.title = element_blank(),                          
        # removes legend title
        legend.key.size = unit(3, "mm"),                         
        # adjusts width of legend color symbols
        legend.key.height = unit(3, "mm"),                       
        # adjusts height of legend color symbols 
        legend.key.spacing.y = unit(1, "mm"),                    
        # defines space between legend items
        panel.grid.minor = element_blank(),                      
        # removes all minor grid-lines 
        panel.grid.major.x = element_blank()) +                  
        # removes grid-lines along the x-axis
  
  labs(x = "Region",                                             
        # x-axis title
       y = "Number of Stomachs Processed",)                      
        # y-axis title

```



This section will include Figure 5  

<!--
Figure 5: 4-paneled bar chart of number of stomach's sampled per size class within each depth stratum. Each panel shows different fish species. 

Data I will need is fish length and the # fish caught per depth stratum per species
-->
```{r stomach-length, echo=FALSE, fig.height = 9.7, fig.width = 7, fig.cap="*Number of stomachs sampled from (a) Atlantic Cod (*Gadus morhua*), (b) Greenland Halibut (*Reinhardtius hippoglossides*), (c) redfish (*Sebastes sp.*), and (d) skate (*Rajidae*) per size class within each depth stratum. Total length was used to measure all predators.*"}
# The purpose of this code chunk is to create Figure 5

# The sections of this code chunk are:
    # Creating and Formatting Figure 5 paneled-bar chart

# No creation of a new dataframe is required for Fig 5


# This figure will detail the total sum of stomachs sampled per size class within each depth stratum.
    # Different panel for each predator category

# Figure structure will need:
    # Three categorical variables:
        # x-axis: predator length
        # fill: depth range
        # facet: predator category
    # One numerical variable:
        # y-axis: sum stomach samples that falls within categories


# If desired, old/test code and additional links to tutorial youtubes/webpages I used to create Figure 5 are recorded in (scripts/draft code/Draft_Script_Fig5_Stomach-by-Length.R)

# Those notes include sites and tutorials that assisted me with:
    # Data structure for figure
    # Building faceted stacked bar charts
    # Formatting faceted stacked bar charts
    # Unused but Interesting ggplot2 arguments
    # Creating a Base Theme for R Markdown Plots


############### Building and Formatting the Grouped Bar Chart ##################


### Faceted Stacked Bar Chart

ggplot(pred, aes(reorder(length.range, Length))) +  
  # orders the X-axis categories based on numerical length of fish
  # basically makes ggplot see and order the x-axis as numbers vs text
  
  geom_bar(aes(fill = depth),
           position = "stack",
           stat="count",
           width = 0.4) +   # changes the width of the bars 
  
  facet_wrap(~ pred.name, 
             nrow = 5, 
             scale = "free_y",
             axes = "all_x") +
  
  theme_minimal() + 
  
  theme(axis.text.x = element_text(angle = 45,hjust = 1, size = 8 ),
        axis.title = element_text(size = 8),
        axis.title.y = element_text(vjust = +3),                 # pulls y-axis title away from chart
        axis.title.x = element_text(vjust = -2),                 # pulls x-axis title away from chart
        legend.title = element_blank(),                          # removes legend title
        legend.key.size = unit(3, "mm"),                         # adjusts width of legend color symbols
        legend.key.height = unit(3, "mm"),                       # adjusts height of legend color symbols
        legend.key.spacing.y = unit(1, "mm"),                    # defines space between legend items
        panel.grid.minor = element_blank(),                      # removes all minor grid-lines
        panel.grid.major.x = element_blank(),
        panel.spacing = unit(13, 'point'),                         # adds space between panels
        strip.text = element_text(face = "bold",      # bolds the panel titles
                                  hjust = 0)) +       # pulls panel titles to the top left
  
  labs(x = "Predator Length (cm)",
       y = "Number of Stomachs")
```


## Proportion of Full vs. Empty Stomachs

This section will include Table 2
<!--
Table 2: Table showing total fish with full and empty stomachs for each fish class size for each fish species

Data I will need includes length (cm) of caught fish, if stomach was full or empty, for each fish species.
-->

```{r table-stomachs, echo=FALSE, fig.cap="*Number of full1, empty, and total (full and empty) stomachs collected for Atlantic Cod (*Gadus morhua*), Greenland Halibut (*Reinhardtius hippoglossides*), redfish (*Sebastes sp.*), and skates (*Rajidae*) within each size class2*"}

```


## Composition of Prey Items

### Composition of Prey Items by %W

### Composition of Prey Items by %N

This section will include Table 3  

<!--
Table 3: Table showing percent weight and percent number of each prey species taxa within the Atlantic cod stomachs. 

Data I will need includes weight and number of prey species within Atlantic cod stomachs
-->

This section will include Table 4  

<!--
Table 4: Table showing percent weight and percent number of each prey species taxa within the Greenland Halibut stomachs. 

Data I will need includes weight and number of prey species within Atlantic cod stomachs
-->

This section will include Table 5

<!--
Table 5: Table showing percent weight and percent number of each prey species taxa within the redfish stomachs. 

Data I will need includes weight and number of prey species within redfish stomachs
-->

This section will include Table 6  

<!--
Table 6: Table showing percent weight and percent number of each prey species taxa within the skates stomachs. 

Data I will need includes weight and number of prey species within skates stomachs
-->

## Pandalus as Prey

### Relative Proportion of Pandalid Prey by Predatory Species

This section will include Figure 6

<!--
Figure 6: 2-paneled bar chart of percent weight and percent number of different shrimp species within the stomachs of the four fish species.  

Data I will need is weight and number of shrimp in each target fish species
-->

### Relative Proportion of Pandalid Prey by Survey Area and Depth

This section will include Figure 7

<!--
Figure 7: 2-paneled bar chart of percent weight and percent number of pandalus shrimp species (combined) within all stomachs by the different depth stratum by fishing area  

Data I will need is weight and number of pandalus shrimp, depth stratum fish was caught, and fishing region that fish were caught
-->

### Relative Proportion of Pandalid Prey by Size Class for Atlantic Cod (*Gadus morhua*) 

This section will include Figure 8

<!--
Figure 8: 2-paneled bar chart of percent weight and percent number different shrimp species within Atlantic cod 

Data I will need is weight and number of shrimp species in Atlantic cod stomachs. 
-->

### Relative Proportion of Pandalid Prey by Size Class for Greenland Halibut (*Reinhardtius hippoglossides*)

This section will include Figure 9

<!--
Figure 9: 2-paneled bar chart of percent weight and percent number different shrimp species within Greenland halibut 

Data I will need is weight and number of shrimp species in Greenland halibut stomachs. 
-->

### Relative Proportion of Pandalid Prey by Size Class for Redfishes (*Sebastes sp.*)

### Relative Proportion of Pandalid Prey by Size Class for Skates (*Rajidae*)

This section will include Figure 10

<!--
Figure 10: 2-paneled bar chart of percent weight and percent number different shrimp species within skate

Data I will need is weight and number of shrimp species in skate stomachs. 
-->

```{r df-exports, include=FALSE}

# This section will export the dataframes you have created.

# !! Make sure you update the years appropriately or old data will be written over !!

write.csv(prey, file="data/processed/2019rm_basePrey.csv")
    # base prey dataframe

write.csv(pred, file="data/processed/2019rm_basePred.csv")
    # base predator dataframe

write.csv(pred.total, file="data/processed/2019_F3_pred.total.csv")
    # Fig 3 dataframe

write.csv(region.total, file="data/processed/2019rm_F4_region.total.csv")
    # Fig 4 dataframe


```


# Acknowledgements

# References

\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent

<!--- 
IN PDF - Creates hanging indent for references. Will not adjust Word doc formatting

Basically instructing text to go back 0.25in on first line, then go in 0.25in on the second
--->


<div id="refs"></div>   

<!--- 
IN PDF - tells program to insert references here. Will not adjust Word doc formatting

NOTE - Waiting for Library to help guide me to correct reference formatting package. Currently testing with a random journal reference csl file
--->

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

<!--- 
Resets paragraph indentation structure in case we have subsequent text
---> 




