---
title: "2019 Pandalus Prey"
author: "Krista Kenyon"
date: "`r Sys.Date()`"
output:
  bookdown::word_document2:
    global_numbering: true
    fig_caption: true
  
---

<!---
Comments for the YAML header above (I can't place this comment above it)

If I want to customize settings of a component (e.g. output), place a ':' at the end.
  I couldn't realize why, say 'toc: true' kept producing errors despite being spaced
  appropriately. It turned out, I did not include the ':' after word_document2:

bookdown::word_document2 = produce a word document AND allow me to cross reference
  figures, tables, and headings.They are only hyperlinked in pdf and html outputs (not word)

NOT NEEDED WITH BOOKDOWN -> number_sections: true = automatically numbering the sections.

global_numbering: true -> word and html outputs label Figures as Fig 1, 2, etc and sections as 1.0, 1.1 etc. This automatically happens with pdf outputs but not the other two.

** At end add in -> 'toc: true'. This adds a  table of contents at the beginning of the document

fig_caption: true = render figures with captions

One user also used 'bibliography' (main function like 'output') to insert which 
  files he had his citations stored. He included 2 files and could reference them.
  (didn't show how). https://www.youtube.com/watch?v=YlcK9Ef31KE

?rmarkdown::word_document= = shows me all the formatting options 

***** I would LOVE to learn how to direct where the output documents are saved.
  I know there are ways.... I just can't find them right now
  
***** For formating pdf text size, fonts, line spacing or insert line numbers etc look at https://www.youtube.com/watch?v=kC6JFcM5e9M
--->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 

# This code chunk sets the parameters for the document code chunks

# The 'include=FALSE' parameter means this particular r chunk won't be shown in the report


######################### Required Packages ####################################


library(ggplot2)
library(tidyverse)
library(dplyr)
library(ggrepel)


################# Notes on cross-referencing figures and tables ################


# In the YAML Header, the 'bookdown::word_document2' parameter gives the ability to cross-reference

# I can reference in text using Figure \@ref(fig:label). It hyperlinks it in pdf but not word. It also gives different Figure number in pdf and word. So there are still things to figure out.
  # to cross reference the figure needs a unique code-chunk name, and have a figure caption

# Code chunk names should be one word or separated with '-', not with spaces, '_', or '.'
  #{r code-chunk}

# I can create figure captions by including 'fig.cap=' in the {r code-chunk, fig.cap=caption text}.

  # If text \@ref(fig:label) is included in notes in the R markdown portion, there will be an error that occurs when Knitting. Therefore included here.
```

<!--
This will form the template for the annual reports presenting the raw fish stomach content species data.
-->

<!--
General formatting notes unique to Rmarkdown.

- If you want text to be on separate lines, you NEED to include 2 spaces at the end of the proceeding line.

- There needs to be a line in between text and a list in order for the formatting to create the list. This does create a space between them in the output as well. I have not looked up if there is a way to adjust the output spacing.
-->

**The following are the pre-amble pages that will need to be included at some point.**

- Document Title Page  <!--- look here https://www.youtube.com/watch?v=kC6JFcM5e9M --->
- Canadian Data Report of Fisheries and Aquatic Sciences (smaller heading)  
- Report Title Page  
- Page of Ownership and How to Cite Report  
- Table of Contents  
- List of Tables  
- List of Figures   

One way to create the Table of Contents is the 'settings function' to the right of 'Knit' -> Output Options -> check Include TOC. This will need to be formatted though to fit the style of the DFO documents.  

<!--- \newpage creates a page break --->
\newpage


# Abstract

# Introduction

This section will include Figure 1
<!--
Figure 1: jpeg of different fishing zones
-->


# Methods

This section will include Figure 2 
<!--
Figure 2: jpeg of the different stomach sampling locations. Looks like made in ArcGIS
-->

This section will include Table 1
<!--
Table 1: # of stomach sampling locations in each zone per depth stratum.

Data that I will need include sample location depths in each stratum.
-->


## Collection of Stomach Samples

## Stomach Content Analysis

## Data Analysis

# Results

```{r import_data, include=FALSE}

# This code chunk:
    # imports the data
    # manipulates the data it to create the base dataframe that will be used throughout the document
    # checks the dataframe structure

# The data imported must be structured consistently in order to utilize this coding.

# All code to check dataframe structure are commented out. If not they will produce outputs into the report.


############################ Key Data Structure Assumption #####################

# It is assumed that the imported data will initially have a structure consistent to the example saved immediately below. Particularly the variable names and variable data type (e.g. chr, int, num)

### Example dataset structure 

# 'data.frame':	2300 obs. of  16 variables:
#  $ content.key: chr  "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" ...
#  $ prey.sp    : int  9998 9998 9998 9998 4770 4770 4770 4770 9998 9998 ...
#  $ weight     : num  0 0 0 0 0.0015 0.0015 0.0015 0.0015 0 0 ...
#  $ count      : int  0 0 0 0 1 1 1 1 0 0 ...
#  $ pred.sp    : int  892 892 892 892 892 892 892 892 892 892 ...
#  $ length     : int  40 40 40 40 45 45 45 45 52 52 ...
#  $ year       : int  2013 2013 2013 2013 2013 2013 2013 2013 2013 2013 ...
#  $ start.depth: int  823 823 823 823 823 823 823 823 823 823 ...
#  $ end.depth  : int  829 829 829 829 829 829 829 829 829 829 ...
#  $ lat.deg    : int  64 64 64 64 64 64 64 64 64 64 ...
#  $ lat.min    : num  27.9 27.9 27.9 27.9 27.9 ...
#  $ long.deg   : int  57 57 57 57 57 57 57 57 57 57 ...
#  $ long.min   : num  59.8 59.8 59.8 59.8 59.8 ...
#  $ sa.code    : chr  "2G" "RISA" "SFA2EX" "SFA3" ...
#  $ sort.key   : int  10 7 6 8 10 7 6 8 10 7 ...
#  $ trawl.id   : int  72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 ...


####################### Importing Your Data ####################################


### import data

stomach <- read.csv('data/raw/2013_test_stomach_data.csv') 
  # assumes header=TRUE and knows it's separated by commas

# str(stomach)
  # check initial data structure. Does it match example?


################### Transforming Data into Base Dataframe ######################

# Additional data variables will be needed in the base dataframe for future figures tables including:
    # column containing count of stomachs sampled per fish. 
    # column with predator 'category' names
        # Categories: Atlantic cod, Greenland halibut, Redfish, Skate


### Adding number stomach sampled per fish

stomach$stomach.count <- 1 


### Adding predator category

stomach$pred.name = stomach$pred.sp  
  # creates new column that is duplicate of pred.sp, and renames to pred. name


stomach <- stomach %>% 
  mutate(pred.name=recode(pred.name, "438"="Atlantic cod", "892"="Greenland halibut", "793" ="Redfish","794"="Redfish", "90"="Skate"))  
  # replaces species codes with predatory categories in new column. 
  # currently the above code only includes the species codes for 2013


### Possible species codes for predator species categories
  # Greenland halibut = 892
  # Atlantic cod = 438
  # Redfish = 792, 793, 794, 997, 998
  # Skates = 80, 88, 89, 90, 91, 92, 94, 95, 96, 97

# See 'Predator Codes and Names.txt' file in 'Scripts' folder for list of species names attached to each code


################### Checking our Base Dataframe Modifications ##################

which(is.na(stomach))
  # checking if there are blank values or NA's in the dataframe

str(stomach) 
  # checking new dataframe structure to ensure all updates are good


### Saving Structure of Base Dataframe

# 'data.frame':	2300 obs. of  18 variables:
#  $ content.key  : chr  "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" "72138002_892_1_1_-171_212" ...
#  $ prey.sp      : int  9998 9998 9998 9998 4770 4770 4770 4770 9998 9998 ...
#  $ weight       : num  0 0 0 0 0.0015 0.0015 0.0015 0.0015 0 0 ...
#  $ count        : int  0 0 0 0 1 1 1 1 0 0 ...
#  $ pred.sp      : int  892 892 892 892 892 892 892 892 892 892 ...
#  $ length       : int  40 40 40 40 45 45 45 45 52 52 ...
#  $ year         : int  2013 2013 2013 2013 2013 2013 2013 2013 2013 2013 ...
#  $ start.depth  : int  823 823 823 823 823 823 823 823 823 823 ...
#  $ end.depth    : int  829 829 829 829 829 829 829 829 829 829 ...
#  $ lat.deg      : int  64 64 64 64 64 64 64 64 64 64 ...
#  $ lat.min      : num  27.9 27.9 27.9 27.9 27.9 ...
#  $ long.deg     : int  57 57 57 57 57 57 57 57 57 57 ...
#  $ long.min     : num  59.8 59.8 59.8 59.8 59.8 ...
#  $ sa.code      : chr  "2G" "RISA" "SFA2EX" "SFA3" ...
#  $ sort.key     : int  10 7 6 8 10 7 6 8 10 7 ...
#  $ trawl.id     : int  72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 72138002 ...
#  $ count.stomach: num  1 1 1 1 1 1 1 1 1 1 ...
#  $ pred.name    : chr  "Greenland halibut" "Greenland halibut" "Greenland halibut" "Greenland halibut" 
```


## Predator Distribution by Assessment Area, Length, and Depth

This section will include Figure 3. 

It will contain cool text that will reference Figure \@ref(fig:pie-chart)
<!--
Figure 3: pie chart showing proportion of fish species caught

Data I will need is simply # of fish caught per species
-->

```{r pie-chart, echo=FALSE, fig.cap="*Proportion of Atlantic Cod (*Gadus morhua*), Greenland Halibut (*Reinhardtius hippoglossides*), redfish (*Sebastes sp.*), and skate (*Rajidae*) stomachs analyzed, calculated as a percent total.*"}

# The purpose of this code chunk is to create Figure 3

# The sections of this code chunk are:
    # Creating a new dataframe for Figure 3
    # Creating and Formatting Figure 3 pie chart
    # Recording Additional Geom_Label arguments that we may want to include in further edits 

# Tutorial and experimental codes for Fig 3 are saved in (scripts/draft code/Draft_Script_Fig3_PieChart.R). This includes links to webpages/youtube

# Additional notes on geom_label and geom_label_repel are saved at the bottom of (Script_Figure3_PieChart.R)

#################### Creating new dataframe for Figure #########################

# This dataframe will detail the total sum of stomachs sampled per predator category

# All code to check dataframe structure are commented out. If not they will produce outputs into the report.


### create figure dataframe

pred.total <- aggregate(stomach$stomach.count, list(stomach$pred.name), sum)

# pred.total  


### Renaming the columns in the new dataframe

names(pred.total)[names(pred.total) == "Group.1"] <- "pred.name"

names(pred.total)[names(pred.total) == "x"] <- "stomach.total"

# names(pred.total)


### Creating figure labels

pred.total$label <- paste(pred.total$stomach.total, sep = "
", 
                          paste0( "(", round(100 * pred.total$stomach.total / sum(pred.total$stomach.total), 0 ), "%)" ))
    # Label structure will be 'total stomach sampled (% of stomachs)'
    # DO NOT ADJUST THE SPACING AFTER [sep = "..."]. This splits the values into 2 lines. Adding spaces will uncenter second line. 
    # paste() puts spaces between labels
    # paste0() does not have spaces between labels. Which means there is no space between # and % (e.g. 40% vs 40 %)


### Checking Final Database Structure

# str(pred.total)


### Saving Final Database Structure 

# 'data.frame':	4 obs. of  3 variables:
#  $ pred.name    : chr  "Atlantic cod" "Greenland halibut" "Redfish" "Skate"
#  $ stomach.total: num  20 1340 900 40
#  $ label        : chr  "20\n(1%)" "1340\n(58%)" "900\n(39%)" "40\n(2%)"


################### Building and Formatting the Pie Chart ######################

# library(ggrepel)   # formats figure labels so they don't overlap


### Creating the pie chart

ggplot(pred.total, aes(x = "", y = stomach.total, fill = pred.name)) +
  coord_polar(theta = "y") + 
        # basing the circular coordinates on the y value (stomach.total)
  
  geom_col(color = "white", linewidth = 0.5) + 
       # creates white line separations between pieces
  
  theme_void() + 
        # removes the default theme (grey background & external grid lines)
  
  geom_label(aes(x = 1.65, label = label),  
        # x=# moves labels along 'x-axis'. Goal here is outside the pie chart.
        # label = pred.total$label created above. 
        # with 'aes()' we don't need 'pred.total$..'. Knows we are in 
    size = 2.5, 
        # adjusts size of label text
    position = position_stack(vjus = 0.5), 
        # defines 'y-axis' position of label. 'vjus=0.5' is centered.
    show.legend = FALSE) +
        # Removes geom_label default of adding an 'a' to legend colors.
  
  theme(
    legend.title = element_blank(), 
        # removes legend title
    legend.key.size = unit(3, "mm"), 
        # adjusts width of legend color symbols
    legend.key.height = unit(3, "mm"), 
        # adjusts height of legend color symbols
    legend.key.spacing.y = unit(1, "mm"),
        # defines space between legend items
  )


####################### Additional geom_label arguments ########################

# These particular arguments are used to adjust features included within the current code parameters. Depending on future formatting, they may be helpful.

# fill = "white",                       
    # changes label fill colour. Default colour matches pie chart chunk colours

# label.padding = unit(1.2, "mm"),       
    # adjusts space between label name and boarder

# label.size =  nit(0.2, "mm"),          
    # adjusts thickness of label boarders

# label.r = unit(3, "mm"),               
    # adjusts roundness of label corners

# family = "montserrat",                
    # this argument would change the font
    # "montserrat" being a font style (would need to import before chart)

```


This section will include Figure 4 
<!--
Figure 4: bar chart of fish caught per region.

Data I will need is # fish caught per species per region
-->

This section will include Figure 5  

<!--
Figure 5: 4-paneled bar chart of number of stomach's sampled from different fish caught at each depth stratum. Each panel shows different fish species. 

Data I will need is # fish caught per depth stratum per species
-->

## Proportion of Full vs. Empty Stomachs

This section will include Table 2
<!--
Table 2: Table showing total fish with full and empty stomachs for each fish class size for each fish species

Data I will need includes length (cm) of caught fish, if stomach was full or empty, for each fish species.
-->

## Composition of Prey Items

### Composition of Prey Items by %W

### Composition of Prey Items by %N

This section will include Table 3  

<!--
Table 3: Table showing percent weight and percent number of each prey species taxa within the Atlantic cod stomachs. 

Data I will need includes weight and number of prey species within Atlantic cod stomachs
-->

This section will include Table 4  

<!--
Table 4: Table showing percent weight and percent number of each prey species taxa within the Greenland Halibut stomachs. 

Data I will need includes weight and number of prey species within Atlantic cod stomachs
-->

This section will include Table 5

<!--
Table 5: Table showing percent weight and percent number of each prey species taxa within the redfish stomachs. 

Data I will need includes weight and number of prey species within redfish stomachs
-->

This section will include Table 6  

<!--
Table 6: Table showing percent weight and percent number of each prey species taxa within the skates stomachs. 

Data I will need includes weight and number of prey species within skates stomachs
-->

## Pandalus as Prey

### Relative Proportion of Pandalid Prey by Predatory Species

This section will include Figure 6

<!--
Figure 6: 2-paneled bar chart of percent weight and percent number of different shrimp species within the stomachs of the four fish species.  

Data I will need is weight and number of shrimp in each target fish species
-->

### Relative Proportion of Pandalid Prey by Survey Area and Depth

This section will include Figure 7

<!--
Figure 7: 2-paneled bar chart of percent weight and percent number of pandalus shrimp species (combined) within all stomachs by the different depth stratum by fishing area  

Data I will need is weight and number of pandalus shrimp, depth stratum fish was caught, and fishing region that fish were caught
-->

### Relative Proportion of Pandalid Prey by Size Class for Atlantic Cod (*Gadus morhua*) 

This section will include Figure 8

<!--
Figure 8: 2-paneled bar chart of percent weight and percent number different shrimp species within Atlantic cod 

Data I will need is weight and number of shrimp species in Atlantic cod stomachs. 
-->

### Relative Proportion of Pandalid Prey by Size Class for Greenland Halibut (*Reinhardtius hippoglossides*)

This section will include Figure 9

<!--
Figure 9: 2-paneled bar chart of percent weight and percent number different shrimp species within Greenland halibut 

Data I will need is weight and number of shrimp species in Greenland halibut stomachs. 
-->

### Relative Proportion of Pandalid Prey by Size Class for Redfishes (*Sebastes sp.*)

### Relative Proportion of Pandalid Prey by Size Class for Skates (*Rajidae*)

This section will include Figure 10

<!--
Figure 10: 2-paneled bar chart of percent weight and percent number different shrimp species within skate

Data I will need is weight and number of shrimp species in skate stomachs. 
-->

# Acknowledgements

# References
